### 2012-11-17 inst/examples/question.Rmd
Illustrate the direct approach to calculating the covariance matrix from the kernlab package.

```{r}
obs <- data.frame(x = c(-4, -3, -1,  0,  2),
                  y = c(-2,  0,  1,  2, -1))
x_predict <- seq(-5,5,len=50)
```

@@

```{r}
sigma.n <- 0.3
cov_xx_inv <- solve(cov(obs$x, obs$x) + sigma.n^2 * diag(1, length(obs$x)))
Ef <- cov(x_predict, obs$x) %*% cov_xx_inv %*% obs$y
Cf <- cov(x_predict, x_predict) - cov(x_predict, obs$x)  %*% cov_xx_inv %*% cov(obs$x, x_predict)
```

@@

```{r}
library(kernlab)
gp <- gausspr(obs$x, obs$y, kernel="rbfdot", kpar=list(sigma=1/(2*l^2)), fit=FALSE, scaled=FALSE, var=0.8)
Ef_kernlab <- predict(gp, x_predict)
```

but I don't see how I get the associated covariances?  Perhaps I have missed something in the documentation?

There are many cases where it would be nice to have access to the resulting Gaussian process, such as in generating plots as in Rasmussen and Williams:

```{r}
require(ggplot2)
dat <- data.frame(x=x_predict, y=(Ef), ymin=(Ef-2*sqrt(diag(Cf))), ymax=(Ef+2*sqrt(diag(Cf))))    
ggplot(dat) +
  geom_ribbon(aes(x=x,y=y, ymin=ymin, ymax=ymax), fill="grey80") + # Var
  geom_line(aes(x=x,y=y), size=1) + #MEAN
  geom_point(data=obs,aes(x=x,y=y)) +  #OBSERVED DATA
  scale_y_continuous(lim=c(-3,3), name="output, f(x)") + xlab("input, x") 
```

>>>>

```{r}
obs <- data.frame(x = c(-4, -3, -1,  0,  2),
                  y = c(-2,  0,  1,  2, -1))
X <- seq(-5,5,len=50)
```

```{r results="hide"}
knit("beverton_holt_data.Rmd")
```

@@

```{r}
sigma_n <- 0.3
cov_xx_inv <- solve(cov(obs$x, obs$x) + sigma_n^2 * diag(1, length(obs$x)))
Ef <- cov(X, obs$x) %*% cov_xx_inv %*% obs$y
system.time(Cf <- cov(X, X) - cov(X, obs$x)  %*% cov_xx_inv %*% cov(obs$x, X))
```

Plot

```{r}
require(ggplot2)
dat <- data.frame(x=X, y=(Ef), ymin=(Ef-2*sqrt(diag(Cf))), ymax=(Ef+2*sqrt(diag(Cf))))    
ggplot(dat) +
  geom_ribbon(aes(x=x,y=y, ymin=ymin, ymax=ymax), fill="grey80") + # Var
  geom_line(aes(x=x,y=y), size=1) + #MEAN
  geom_point(data=obs,aes(x=x,y=y)) +  #OBSERVED DATA
  scale_y_continuous(name="output, f(x)") + xlab("input, x") 
```

@@

rename hyperparameters

```{r}
l = 1
lengthscale = 1/(2*l^2)
var = sigma_n^2
```

```{r}
library(kernlab)
#gp <- gausspr(obs$x, obs$y, kernel="rbfdot", kpar=list(sigma=0.5), fit=FALSE, scaled=FALSE, var=.09)
gp <- gausspr(obs$x, obs$y, kernel="rbfdot", scaled=FALSE, var=.09)
Ef_k <- predict(gp, X)
```

Manually get the covariance.  First we compute $K(x, x)$

```{r}
K <- kernelMatrix(kernelf(gp), xmatrix(gp))
```

Then compute $(K(x,x) + \sigma_n^2 \mathbb{I})^{-1}$,

```{r}
Inv <- solve(K + diag(rep(gp@kcall$var, length = length(xmatrix(gp)))))
```

The covariance is $K(x_*, x_*) - K(x_*, x)(K(x, x) - \sigma_n^2\mathbb{I})^{-1}K(x,x_*)$.

```{r}
system.time(Cf_k <- kernelMatrix(kernelf(gp), as.matrix(X)) - kernelMatrix(kernelf(gp), as.matrix(X), xmatrix(gp)) %*% Inv %*% kernelMatrix(kernelf(gp), xmatrix(gp), as.matrix(X)))
```

Identical to computing the $K(x_*, x)(K(x, x) - \sigma_n^2\mathbb{I})^{-1}$ part first, multiplying the off-diagonal matrix block by the inverse

```{r}
system.time(Cf_k2 <- kernelMatrix(kernelf(gp), as.matrix(X)) - kernelMult(kernelf(gp), as.matrix(X), xmatrix(gp),  Inv) %*% kernelMatrix(kernelf(gp), xmatrix(gp), as.matrix(X)))
```

which is also faster.

compare the two in plots

```{r}
require(ggplot2)
dat_k <- data.frame(x=X, y=(Ef_k), ymin=(Ef_k-2*sqrt(diag(Cf_k))), ymax=(Ef_k+2*sqrt(diag(Cf_k))))    
ggplot(dat_k) +
  geom_ribbon(aes(x=x,y=y, ymin=ymin, ymax=ymax), fill=rgb(0,0,1,.5)) + # Var
  geom_line(aes(x=x,y=y), size=1, col="blue") + #MEAN
  geom_point(data=obs,aes(x=x,y=y)) +  #OBSERVED DATA
  geom_ribbon(data=dat, aes(x=x,y=y, ymin=ymin, ymax=ymax), fill=rgb(1,0,0,.5)) + # Var
  geom_line(data=dat, aes(x=x,y=y), size=1, col="red") + #MEAN
  scale_y_continuous(name="output, f(x)") + xlab("input, x") 
```
